Notes on conformal manifolds

### 2025-09-22 Draft_MLFlatDirections.tex
Updated section 3 and added a few minor edits on sec 4

MI-HET-??? \\
\today
\end{flushright}

@@

\begin{center}
\textbf{Abstract}
\end{center}

\begin{quote}

...

\end{quote}

@@

\paragraph{}
The paper is organized as follows. In sec.~\ref{sec:sugra}, we establish the supergravity setup, presenting the scalar potential in its full complexity and motivating the restriction to five fields. Sec.~\ref{Sec:AIS-SMC} details our annealed importance sampling approach to symbolic regression. We present in sec.~\ref{sec:results} our main results, both numerical an analytical. Sec.~\ref{sec:sugrasol} gives some details on the new supergravity solutions discovered by the numerical analysis. Finally, sec.~\ref{sec:ccl} concludes with prospects for extending this approach to higher-dimensional cases and its broader implications for systematic studies of conformal manifolds in holographic theories.

@@

The numerical search described above results in clouds of points embedded in a higher-dimensional space. 
In the realm of symbolic regression, the aim is to uncover interpretable mathematical expressions that best describe such datasets. In the problem discussed in this paper, we are looking for polynomial expressions such that $p(\vec{x}^{\,i}) = 0$, with $i \in \{1, \dots, \texttt{n\_{sample}}\}$. This task involves navigating a vast, discrete, and often rugged search space of possible symbolic models, which poses significant challenges for traditional sampling methods. Markov Chain Monte Carlo (MCMC) techniques such as Metropolis-Hastings or Gibbs sampling, while widely used, can struggle with poor mixing and getting trapped in local optima, especially in high-dimensional or multimodal spaces.

To address these challenges, we employ an Annealed Importance Sampling (AIS) \cite{neal1998annealedimportancesampling} combined with Sequential Monte Carlo (SMC) methods. 
AIS aims to estimate ratios of partition functions by constructing a sequence of intermediate distributions that smoothly transition from an initial, tractable distribution (e.g., a prior over symbolic expressions) to the complex target posterior distribution. This annealing process is guided by a temperature-like parameter that gradually emphasises the data likelihood, allowing for more efficient exploration of the probability landscape.

SMC enhances this procedure by propagating a population of particles—each representing a candidate symbolic expression—through the sequence of distributions. At each step, particles are reweighted based on the incremental change in the distribution, resampled to focus computational effort on high-probability regions, and mutated via operations. This combination of importance sampling, resampling, and mutation maintains diversity among the particles and prevents premature convergence to suboptimal models.
These features make AIS-SMC particularly well-suited for symbolic regression tasks, where the search space is not only high-dimensional but also structured and discontinuous.

@@

Let us now explain in more detail how does this procedure goes. The goal is to reconstruct some distribution function $d(z)$, where here $z$ in going to be some polynomial. We will try to reconstruct this density function by series of density function $\pi_n(z_n) = \gamma_n(z_n)/Z_n$ with $n = 1,\dots,N$ the number of annealing steps, $\pi_n$ is defined in terms of an unnormalized density $\gamma_n$ and we have the normalizing constant $Z_n = \int \gamma_n(z) \mathrm{d}z$. We also assume we have a sequence of inverse temperature constants $\beta_n$, where $0 = \beta_1 < \beta_2 <\dots< \beta_N$. The unnormalized density at level $n$ is then given in terms of a prior distribution $p_0(z)$ over the hypothesis space and a loss function $L(z)$:
		\begin{equation}\label{eq:gamman}
			\gamma_n(z) := d_0(z) \: \exp \left( -\beta_n L(z) \right).
		\end{equation}

		At each epoch, we have a set of particles and weights $\{z_{n-1}^k, w_{n-1}^k\}$ approximating $\pi_{n-1}$ (meaning $\mathbb{E}_{\pi_{n-1}}[f] \approx \frac{\sum_k w_{n-1}^k f(z_{n-1}^k)}{\sum_k w_{n-1}^k}$), and we want to obtain a new set $\{z_n^k, w_n^k\}$ approximating $\pi_n$. To do so, for each particle $z^k_{n-1}$, we propose a new particle $z_n^k \sim q(z_n | z_{n-1}^k)$ and calculate the new unnormalized importance weight $w_n^k$. The latter are updated using the formula 
		\begin{equation}
			w_n^k = w_{n-1}^k \times \alpha_{n}^k
		\end{equation}
		where $\alpha_n^k$ is the incremental importance weight. The standard form for $\alpha_n^k$, which requires introducing an auxiliary backward transition kernel $q(z_{n-1} | z_n)$, is:
		\begin{equation}
			\alpha_n^k = \frac{ \gamma_n(z_n^k) q(z_{n-1}^k | z_n^k) }{ \gamma_{n-1}(z_{n-1}^k) q(z_n^k | z_{n-1}^k) } 
		\end{equation}
		The weights are finally normalized and we get $w_n^k \rightarrow \frac{w_n^k}{\sum_j w_n^j}$.

		Let's us now focus on the forward $q(z_n^k | z_{n-1}^k)$ and backward propagation kernel $q(z_{n-1}^k | z_n^k)$ for now. The forward propagation kernel defines how we generate the state at time $n$ given the state at time $n-1$. It gives the probability to transition from $z_{n-1}$ to $z_n$. The backward kernel represents a hypothetical probability of transitioning back from state $z_n$ to state $z_{n-1}$. The purpose of $q$ is to give a proposition for $z_n$ given $z_{n-1}$. For the implementation of the two, we decide to implement an AIS-style MCMC code: we make some move in the space of polynomials, and then accept or reject those new polynomials based on an acceptance rate. So we first need to choose what are the available moves in the space of polynomials. Here are the choices we have made for our symbolic regression task : 
		\begin{itemize}
			\item Coefficient pertubation. Given a polynomial $z_{n-1}$, we choose randomly one of its coefficient and modify it by a random noise from the guassian distribution $\mathcal{N}(0,\sigma^2)$. So for eg : $2 x_1 + 3 x_2^2 \rightarrow  2.1 x_1 + 3 x_2^2$.
			\item Variable multiplication. Given a polynomial $z_{n-1}$, we choose randomly one of his monomial, and multiply it by one of the available variable. So for eg : $2 x_1 + 3 x_2^2 \rightarrow  2 x_1 x_2 + 3 x_2^2$.
			\item Variable division. Given a polynomial $z_{n-1}$, we choose randomly one of his monomial, and divide it by one of its variable. So for eg : $2 x_1 x_2 + 3 x_2^2 \rightarrow  2 x_1 + 3 x_2^2$. 
		\end{itemize}
		The polynomials get mutated at each epoch. The mutation is chosen randomly, assigning probabilities \texttt{p\_{modify}} , \texttt{p\_{multiply}} and \texttt{p\_{divide}} to each possibility. This implementation enables a direct computation of both the forward and backward propagation kernels. In particular, for the case of coefficient perturbations, the transition is symmetric. That is,
		\begin{equation}
			q(z_n^k \mid z_{n-1}^k) = q(z_{n-1}^k \mid z_n^k),
		\end{equation}
		which significantly simplifies the computation of the incremental importance weight.

		However, this symmetry does not hold in general. For other types of operations—such as multiplication or division by a variable—the transition from \( z_{n-1}^k \) to \( z_n^k \) is generally not equivalent to the reverse transition from \( z_n^k \) to \( z_{n-1}^k \). In such cases $q(z_n^k \mid z_{n-1}^k) \ne q(z_{n-1}^k \mid z_n^k)$ and particular care must be taken in the computation of the backward propagation kernel.

		The acceptance ratio for the MCMC algorithm is given by : 
		\begin{equation}\label{eq:acceptanceratio}
			A(z_n,z_{n-1}) = \mathrm{min} \left(1,\frac{ \gamma_n(z_n^k) q(z_{n-1}^k | z_n^k) }{ \gamma_{n-1}(z_{n-1}^k) q(z_n^k | z_{n-1}^k) }  \right)
		\end{equation}
		We then draw $u\sim \mathrm{Uniform}(0,1)$, accept the new particle if $u<A(z_n,z_{n-1})$ and reject it otherwise.

		We have adapted the classic approach outlined above. As can be seen in \eqref{eq:gamman}, what matters the most to determine $\gamma_{n}$ is not the inverse temperature, but rather the product of the inverse temperature and the loss function. Therefore, instead of implementing an arbitrary schedule for $\beta$, we employed an adaptive temperature approach. The number of particles accepted at each run is monitored. We aim at following the following schedule as function of the \texttt{epoch}:
		\begin{equation} \label{eq:eqadaptivetemperatureschedule}
			\texttt{adaptive\_temperature\_ratio} = 0.8 - 0.5 \times \left(\frac{\texttt{epoch}}{\texttt{n\_epochs}}\right),
		\end{equation}
		where \texttt{n\_epochs} is the total number of epochs. If we accept more particle (in the 5 last epochs), we decrease the temperature by 5\%, if we accept less we increase it by 5\%. This encourages exploration during the early stages of training and gradually transitions to a more selective regime. %ICI \bd{Move \eqref{eq:eqadaptivetemperatureschedule} in section 4?}

		Before closing the section, we need to discuss what is the Loss function we choose. It is required to compute the unormalized distribution $\gamma_n$. Assume that we have $z = \sum_k c_k X_k$, where $c_k$ are the coefficients of the polynomials, and $X_k$ is a short notations for all the possible monomials up to a given degree (so if we have $x$ and $y$ as variables, and the maximum degree is 2, then $X_k$ are $1, x, y, x^2, y^2, x y$). For the Loss function, we decided to take : 
		\begin{equation} \label{eq:lossreg}
			L(z) = \sum_{i} z(x_i)^2 + \frac{\lambda}{\sum_k |c_k|}
		\end{equation}
		So the first term is just the sum of the squared of the polynomial evaluated on the data. We want to make this 0 so we find polynomial that annihilates our data. The second part is a regularisation factor: it prevents the algorithm to send all the coefficient to 0, which would give a trivial solution to the problem. We typically take $\lambda \sim \mathcal{O}(1000)$.

		%\item Analysis after AIS: select the best polynomials and do exploitation on the coefficients (without MC: we keep only the proposal of it betters the polynomial).

		Once the Annealing loop is over, we end up with a total of $n_{\rm sample}$ particles, which in principle should be close to annihilate our data, but whose coefficients may need some more fine tuning. To deal with it, we ran a quick exploitation phase, where, for each polynomial, we now only modify its coefficients, with a perturbation $\epsilon \sim \mathcal{N}(0, \sigma_1)$, and keep the new particle, only if the Loss function is now getting smaller. This allow to fine-tune the coefficients. %%ICI \ce{Mention threshold? And number of iterations?}

@@

%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%
\section{Numerical Analysis and Results} \label{sec:results}
We now detail the numerical analysis and present our main results. We start by describing how we sampled the solutions manifold using a gradient descent. Once the manifold correctly sampled, we determine its dimension using a local principal component analysis and a clustering procedure. We finally exploit the annealed importance sampling algorithm described in sec.~\ref{Sec:AIS-SMC} to find an analytical characterisation of the manifold.

	\subsection{Sampling the manifold: gradient descent} \label{sec:graddes}
	To perform the gradient descent, we initialise randomly and uniformly points within a hypercube of range $[-2,2]$. This choice is important to ensure that points are not restricted to the inner range~$[-1,1]$, which will be crucial for the Symbolic Regression.\footnote{The rationale is that symbolic regression (SR) evaluates the data points $\vec{x}^{(i)}$ using a candidate polynomial function $p$. The loss function—used later to assess the quality of a given polynomial—is defined as $\sum_i p(\vec{x}^{(i)})$. When all data points lie within the interval $[-1, 1]$, the SR algorithm tends to favor high-degree polynomials, as these often yield smaller values over this range, thereby artificially lowering the loss. Introducing data points with absolute values greater than 1 mitigates this bias by discouraging the selection of polynomials that exploit this effect.}. As the example we are focussing on has a five-dimensional parameter space, we choose to generate $n_{\rm points}=10^5$ points: we aim at populating all five directions and want approximately $\mathcal{O}(10)$ points per direction. As the solutions lie within the five dimensional space, its intrinsic dimension is less than or equal to 5. The value $10^5$ then serves as a conservative upper bound for the number of points needed to adequately sample the manifold.

@@

	\begin{figure}
		\centering
		\includegraphics[scale = 0.75]{Figures/Loss_grad_des_21_05_2025.pdf}
		\caption{Evolution of the loss function~\eqref{eq:lossgraddes} during gradient descent (right $y$-axis, in blue) and corresponding learning rate schedule $\alpha$ (left $y$-axis in orange), both plotted on a logarithmic scale. Dashed lines indicate epochs at which the optimizer was reinitialized.}
		\label{fig:loss_grad_des}
	\end{figure}

	The Adam optimizer was employed throughout the gradient descent procedure~\cite{kingma2017adammethodstochasticoptimization}. As an accelerator of convergence, we observed that periodically reinitialising the optimizer significantly improved the convergence rate. It was reinitialised at iterations {250, 500, 750}, with the learning rate $\alpha$ fixed at $10^{-2}$. At iteration 1000, the optimizer was reinitialised once more, this time with a reduced learning rate of $10^{-3}$. A final reinitialisation was performed at iteration 1500, setting the learning rate to $10^{-4}$, and the optimization was continued for an additional 500 epochs. The evolution of the loss function~\eqref{eq:lossgraddes}, along with the learning rate schedule, is shown in fig.~\ref{fig:loss_grad_des}. As can be seen in this figure, the convergence rate improves significantly each time the optimizer is rebooted. We also observe that, following the last few reinitialisations, the loss exhibits a small bump immediately after the restart. We interpret this behaviour as follows: the learning rate gets internally adjusted by the Adam optimizer during the descent, and may well be smaller than the instructions when the reset occurs. The learning rate then gets suddenly increased, and some points that previously had low loss values may momentarily worsen before benefiting from faster convergence. This effect of faster convergence after the reinitialization is likely due to points initially located in regions with weak attractive basins being pushed toward areas where the potential gradient is steeper, thus accelerating their convergence. This strategy introduces the risk of desampling certain regions in favour of others, introducing the risk to hide some flat directions in the potential. However, we believe that with enough points, it is very unlikely for a flat direction to be completely hidden from us and desampled. This method bears similarity to the concept of warm restarts introduced in ref.~\cite{loshchilov2017sgdrstochasticgradientdescent}, where the authors reset the learning rate to some value at each reset, without reinitialising the whole optimizer. However, we found that simply scheduling the learning rate without resetting the optimizer yielded slower convergence. We interpret this as follows: when the optimizer is reinitialised, it effectively "forgets" its past gradient history. As a result, the actual learning rate used corresponds more closely to the specified value, rather than being internally adjusted based on accumulated past gradients. This effect seems to contribute to faster convergence in our case.

	\begin{figure}
		\centering
		\subfigure{ \label{fig:log_grad_V}
			\begin{tikzpicture}
		  		\draw (0,0) node (fig1) {\includegraphics[scale=0.7]{Figures/log_grad_V_hist_21_05_2025.pdf}};
		  		\draw ($(fig1.north west) + (0,0)$) node {\small (a)};
		  		\draw (9,0) node (fig2) {\includegraphics[scale=0.7]{Figures/log_Vp4_hist_21_05_2025.pdf}};
		  		\draw ($(fig2.north west) + (0,0)$) node {\small (b)};
		  	\end{tikzpicture}
		}
		\subfigure{\label{fig:log_Vp4}}
		\vspace{-1cm}
		\caption{(a) Histogram of the the norm of the gradient for each point ($x$-axis is in $\log$ scale). (b)~Histogram of the distance of the value of the potential to $V(0)$ for each point ($x$-axis is in $\log$ scale). Note that four points have deviations of the order of $10^{-3}$ and are not visible on the figure.}
		\label{fig:Analyze_V_points}
	\end{figure}

	Upon completion, the loss function converges around $0.002$. The distribution of the values of the gradient at each of the data points is plotted in fig.~\ref{fig:log_grad_V}. More than $93\%$ of the points converged to values of $\vert\nabla V\vert$ lower than $10^{-4}$, and more than $99\%$ of them exhibited gradient norms smaller than $10^{-3}$, yielding a satisfying sampling of the flat directions. As can be observed in fig.~\ref{fig:log_Vp4}, all data points converged to values close to $V(0)=-4$. Specifically, nearly all points attained $V(0)$ within an absolute error of at least $10^{-4}$, with the exception of four points whose deviations were of the order of $10^{-3}$ (not visible on the graph). As the origin $\vec{X}=0$ of the parameter space corresponds to the three-dimensional truncation of the round ${\rm AdS}_{3}\times S^{3}$ solution of half-maximal supergravity in six dimensions with $V(0)=-4$, this confirms that the data points lie within flat directions of the potential.

@@

	As a first visualisation, we present a tomography of the data in fig.~\ref{triangular_plot_1_2_4_8_10}. This figure shows all possible 2d projections of the data, along with the 1d histograms of each coordinate after gradient descent. Note that all directions appear to be well populated. There also seems to be non-trivial correlations in the data, see for example the $x_1/x_2$, $x_4/\tilde{x}_8$ or $x_4/x_{10}$ graphs. They could be genuine correlations, or may result from larger basins of attraction. Thanks to the analysis of sec.~\ref{sec:results}, we will conclude that the structures in the $x_1/x_2$ plot are artefacts of the gradient descent, whereas the ones seen in the $x_4/\tilde{x}_8$ or $x_4/x_{10}$ plots reflect genuine features of the manifold.

>>>>

\today
\end{flushright}

@@

\begin{abstract}
	We investigate the structure of conformal manifolds around AdS$_3 \times S^3$ which lift from continuous flat directions in the scalar potential of gauged supergravity resulting from six-dimensional $\mathcal{N}=(1,1)$.
	Our approach combines numerical exploration and symbolic inference. For the latter, we develop a symbolic regression algorithm based on Sequential Importance Sampling (AIS) combined with Sequential Monte Carlo (SMC), well-suited to uncovering polynomial constraints in high-dimensional parameter spaces. 
	The algorithm reconstructs a set of polynomial relations that provides an explicit analytic parametrization of a new family of solutions. 
\end{abstract}

@@

\paragraph{}
The paper is organized as follows. In sec.~\ref{sec:sugra}, we establish the supergravity setup, presenting the scalar potential in its full complexity and motivating the restriction to five fields. Sec.~\ref{Sec:AIS-SMC} details our annealed importance sampling approach to symbolic regression. Our main results, both numerical an analytical, are presented in sec.~\ref{sec:results}, and sec.~\ref{sec:sugrasol} gives some details on the new supergravity solutions discovered by the numerical analysis. We conclude with prospects for extending this approach to higher-dimensional cases and its broader implications for systematic studies of conformal manifolds in holographic theories.

@@

The numerical sampling of the space of vacua described above results in clouds of points embedded in a higher-dimensional space. 
In the realm of symbolic regression, the aim is to uncover interpretable mathematical expressions that best describe the embedding of these loci of solutions. 
In the present context, the form of the supergravity potential \eqref{eq:scalarpotential} determining the solutions implies that the loci can be defined through polynomial expressions such that $p(\vec{x}^{\,i}) = 0$, with $i \in \{1, \dots, \texttt{n\_{sample}}\}$. 
Finding such polynomials involves navigating a vast, discrete, and often rugged search space of possible symbolic models, which poses significant challenges for traditional sampling methods. 
Markov Chain Monte Carlo (MCMC) techniques such as Metropolis-Hastings or Gibbs sampling, while widely used, can struggle with poor mixing and often get trapped in local optima, especially in high-dimensional or multimodal spaces \cite{??}.

To address these challenges, we employ an Annealed Importance Sampling (AIS) \cite{neal1998annealedimportancesampling} combined with Sequential Monte Carlo (SMC) methods \cite{naesseth2024elementssequentialmontecarlo}. 
AIS aims to estimate ratios of partition functions by constructing a sequence of intermediate distributions that smoothly transition from an initial, tractable distribution (e.g. a prior over symbolic expressions) to the complex target posterior distribution. This annealing process is guided by a temperature-like parameter that gradually emphasises the data likelihood, allowing for more efficient exploration of the probability landscape.

SMC realises this procedure by encoding the probability distributions into a set of particles—each representing a candidate symbolic expression— and weights, and making them evolve so as to focus computational effort on high-probability regions. This combination of importance sampling, resampling, and mutation maintains diversity among the particles and prevents premature convergence to suboptimal models.
These features make AIS-SMC particularly well-suited for symbolic regression tasks, where the search space is not only high-dimensional but also structured and discontinuous.

@@

Let us now explain in more detail the procedure. 
The goal is to obtain a probability distribution $d(z)$, where $z$ runs over a space of polynomials. 
This function assigns high probabilities to those polynomials that annihilate the data, and therefore will be used to identify such symbolic expressions.
We reconstruct this density function by series of density functions $\pi_n(z)$ such that $\pi_n\to d(z)$ as $n\to\infty$. 
On every annealing step, $\pi_n$ is defined in terms of an unnormalised density $\gamma_n$ and a normalisation constant as $\pi_n(z) = \gamma_n(z)/Z_n$.
The unnormalised density at level $n$ is given in terms of a prior distribution $d_0(z)$ over the hypothesis space and a loss function $L(z)$:
		\begin{equation}\label{eq:gamman}
			\gamma_n(z) = d_0(z) \: \exp \left( -\beta_n L(z) \right),
		\end{equation} 
where the inverse temperature constants $\beta_n$ are taken to evolve monotonically as $\beta_1 < \beta_2 <\dots< \beta_N$.

		Sequential importance sampling \cite{...} further assumes that the density functions $\pi_n(z)$ can be given in terms of importance densities $q^k_n(z)$ and weights $w^k_n$ as $\pi_n(z) = \sum w^k_n q^k_n(z)/Z_n$, and these importance densities can be sequentially given as
		 \begin{equation}
		 	q_n(z_n) = q_{n-1}(z_{n-1}) q_{n\vert n-1}(z_n\vert z_{n-1})\,.
		 \end{equation}
At each epoch, we $\pi_{n}$ can be approximated via a set of particles and weights $\{z_{n}^k,\ w_{n}^k\}$  so that\footnote{In the following, we always normalise the weights so that $\sum_k w_{n}^k=1$.}
		 \begin{equation}
		 	\pi_n(z) \approx \sum_k w_{n}^k\ \delta(z-z_{n}^k)\,,
		 \end{equation}
To obtain a new set $\{z_{n+1}^k,\ w_{n+1}^k\}$ approximating $\pi_{n+1}$, we sample from the forward propagation kernel $q_{n+1}(z_{n+1}\vert z_n)$ that defines how we generate the state at time $n+1$ given the state at time $n$.
Similarly, the evolution of the weights is given by
\[
	w_{n+1} = \frac{\gamma_{n+1}}{q_{n+1}} = \frac{\gamma_{n+1}}{q_{n}\ q_{n+1 \vert n}} = w_{n}\ \alpha_{n+1}\,,
\]
where the incremental importance weights $\alpha_{n}^k$ are given by
		\begin{equation}	\label{eq: incremental_importance}
			\begin{aligned}
				\alpha^k_{n+1} &= \frac{\gamma_{n+1}(z_{n+1})}{\gamma_{n}(z_n)}\frac1{q^k_{n+1 \vert n}(z_{n+1}\vert z_{n})}	\\
				&=e^{-\beta_nL(z_n)+\beta_{n+1}L(z_{n+1})}\frac1{q^k_{n+1 \vert n}(z_{n+1}\vert z_{n})}\,.
			\end{aligned}
		\end{equation}

		For the implementation of the forward propagation kernel, we perform an AIS-style MCMC algorithm. 
On every step, we make some move in the space of polynomials, and then accept or reject those new polynomials based on a fixed rate.
The moves that have been allowed in our implementation are:
		\begin{itemize}
			\item \textbf{Coefficient pertubation}. Given a polynomial $z_{n}$, we choose one of its coefficients at random and modify it by a Gaussian noise distributed as $\mathcal{N}(0,\sigma^2)$. E.g.
			\[
				2 x_1 + 3 x_2^2 \mapsto  2.1 x_1 + 3 x_2^2\,.
			\]
			\item \textbf{Variable multiplication}. Given a polynomial $z_{n}$, we pick randomly one of its monomials, and multiply it by one of the available variables. E.g.
			\[
				2 x_1 + 3 x_2^2 \mapsto  2 x_1 x_2 + 3 x_2^2\,.
			\]
			\item \textbf{Variable division}. Given a polynomial $z_{n}$, we choose randomly one of its monomials, and divide it by one of its variables. E.g.
			\[
				2 x_1 x_2 + 3 x_2^2 \mapsto  2 x_1 + 3 x_2^2\,.
			\]
		\end{itemize}
		Each of these operations is chosen randomly, with probabilities $p_{\rm shift}$, $p_{\rm multiply}$, and $p_{\rm divide}$.
%		For the case of coefficient perturbations, the transition is symmetric. 
%		However, this symmetry does not hold in general. For other types of operations—such as multiplication or division by a variable—the transition from \( z_{n-1}^k \) to \( z_n^k \) is generally not equivalent to the reverse transition from \( z_n^k \) to \( z_{n-1}^k \). In such cases $q(z_n^k \mid z_{n-1}^k) \ne q(z_{n-1}^k \mid z_n^k)$ and particular care must be taken in the computation of the backward propagation kernel.
After performing these updates, the change is accepted based on the temperature-dependent rate
		\begin{equation}\label{eq:acceptanceratio}
			A(z_n,z_{n-1}) = \mathrm{min} \left(1,\frac{ \gamma_n(z_n^k) q(z_{n-1}^k | z_n^k) }{ \gamma_{n-1}(z_{n-1}^k) q(z_n^k | z_{n-1}^k) }  \right)\,.
		\end{equation}
		We then draw $u\sim \mathrm{Uniform}(0,1)$, accept the new particle if $u<A(z_n,z_{n-1})$ and reject it otherwise.
		The kernels $q$ and $\tilde{q}$ thus encapsulate the effects of performing these operations on the polynomials at step $n$.

		To compute the weights and kernels via \eqref{eq: incremental_importance} and \eqref{eq:acceptanceratio} we need to consider the loss function in \eqref{eq:gamman}.
In the present context, we use the following loss, where the polynomials have been parameterised as $z = \sum_k c_k X_k$ with $X_k$ denoting the possible monomials up to a given degree,%
\footnote{So, if we have $x$ and $y$ as variables, and the maximum degree is 2, then $X_k$ are $1, x, y, x^2, y^2, x y$.}
		\begin{equation} \label{eq:lossreg}
			L(z) = \sum_{i} z(x_i)^2 + \frac{\lambda}{\sum_k |c_k|}\,.
		\end{equation}
Here, the first term is just the sum of the square of the polynomial evaluated on the data, and is therefore minimal when the polynomial annihilates the data. 
The second term is a regularisation factor that prevents the algorithm to send all the coefficient to 0, which would give a trivial solution to the problem. 
We typically take $\lambda \sim \mathcal{O}(10^3)$. 
Together with the loss function, we also need to specify a cooling schedule for $\beta$. We will discuss different approaches in the next section.

		Once the Annealing loop is over, we end up with a total of $n_{\rm sample}$ particles, which in principle should be close to annihilate our data, but whose coefficients may need some refinement. To deal with it, for each polynomial we now only modify its coefficients with a perturbation $\epsilon \sim \mathcal{N}(0, \sigma')$ and keep the particles when the loss function decreases.

@@

%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%
\section{Numerical Analysis and Results} \label{sec:results}
After describing the theory we want to address in sec.~\ref{sec:sugra}, and the symbolic regression method to be employed in sec.~\ref{Sec:AIS-SMC}, we now proceed to detail the numerical analysis and present our main results. 
We start by describing how we sampled the space of solutions using gradient descent. Once the manifold correctly sampled, we determine its dimension using a local principal component analysis and a clustering procedure. We finally exploit the ASMC algorithm described in sec.~\ref{Sec:AIS-SMC} to find an analytical characterisation of the manifold.

	\subsection{Sampling the manifold: gradient descent} \label{sec:graddes}
	To perform the gradient descent, we initialise randomly and uniformly points within a hypercube of range $[-2,2]$. This choice is important to ensure that points are not restricted to the inner range~$[-1,1]$, which will be crucial for the symbolic regression.\footnote{The rationale is that our symbolic regression evaluates the data points $\vec{x}^{(i)}$ using a candidate polynomial function $z$. When all data points lie within the interval $[-1, 1]$, the algorithm tends to favor high-degree polynomials artificially lowering the loss. Introducing data points with absolute values greater than 1 mitigates this bias.} As the example we are focussing on has a five-dimensional parameter space, we choose to generate $n_{\rm points}=10^5$ points, since we aim at populating all five directions and want approximately $\mathcal{O}(10)$ points per direction. As the solutions lie within the five dimensional space, its intrinsic dimension is less than or equal to 5. The value $10^5$ then serves as a conservative upper bound for the number of points needed to adequately sample the manifold.

@@

	\begin{figure}
		\centering
		\includegraphics[scale = 0.75]{Figures/Loss_grad_des_21_05_2025.pdf}
		\caption{Evolution of the loss function $\mathcal{L}$ in~\eqref{eq:lossgraddes} during gradient descent (right axis, in blue) and corresponding learning rate schedule $\alpha$ (left axis in orange), both plotted in logarithmic scale. Dashed lines indicate epochs at which the optimizer was reinitialized.}
		\label{fig:loss_grad_des}
	\end{figure}

	The Adam optimizer was employed throughout the gradient descent procedure~\cite{kingma2017adammethodstochasticoptimization}. As an accelerator of convergence, we observed that periodically reinitialising the optimizer significantly improved the convergence rate. It was reinitialised at iterations {250, 500, 750}, with the learning rate $\alpha$ fixed at $10^{-2}$. At iteration 1000, the optimizer was reinitialised once more, this time with a reduced learning rate of $10^{-3}$. A final reinitialisation was performed at iteration 1500, setting the learning rate to $10^{-4}$, and the optimization was continued for an additional 500 epochs. The evolution of the loss function~\eqref{eq:lossgraddes}, along with the learning rate schedule, is shown in fig.~\ref{fig:loss_grad_des}. As can be seen in this figure, the convergence rate improves significantly each time the optimizer is rebooted. We also observe that, following the last few reinitialisations, the loss exhibits a small bump immediately after the restart. We interpret this behaviour as follows: the learning rate gets internally adjusted by the Adam optimizer during the descent, and may well be smaller than the instructions when the reset occurs. The learning rate then gets suddenly increased, and some points that previously had low loss values may momentarily worsen before benefiting from faster convergence. This effect of faster convergence after the reinitialization is likely due to points initially located in regions with weak attractive basins being pushed toward areas where the potential gradient is steeper, thus accelerating their convergence. This strategy introduces the risk of desampling certain regions in favour of others, introducing the risk to hide some flat directions in the potential. However, we believe that with enough points, it is very unlikely for a flat direction to be completely desampled. This method bears similarity to the concept of warm restarts introduced in ref.~\cite{loshchilov2017sgdrstochasticgradientdescent}, where the authors reset the learning rate to some value at each reset, without reinitialising the whole optimizer. However, we found that simply scheduling the learning rate without resetting the optimizer yielded slower convergence. We interpret this as follows: when the optimizer is reinitialised, it effectively "forgets" its past gradient history. As a result, the actual learning rate used corresponds more closely to the specified value, rather than being internally adjusted based on accumulated past gradients. This effect seems to contribute to faster convergence in our case.

	\begin{figure}
		\centering
		\subfigure{ \label{fig:log_grad_V}
			\begin{tikzpicture}
		  		\draw (0,0) node (fig1) {\includegraphics[scale=0.7]{Figures/log_grad_V_hist_21_05_2025.pdf}};
		  		\draw ($(fig1.north west) + (0,0)$) node {\small (a)};
		  		\draw (9,0) node (fig2) {\includegraphics[scale=0.7]{Figures/log_Vp4_hist_21_05_2025.pdf}};
		  		\draw ($(fig2.north west) + (0,0)$) node {\small (b)};
		  	\end{tikzpicture}
		}
		\subfigure{\label{fig:log_Vp4}}
		\vspace{-1cm}
		\caption{(a) Histogram of the the norm of the gradient for each point ($x$-axis is in $\log$ scale). (b)~Histogram of the distance of the value of the potential to $V(0)$ for each point ($x$-axis is in $\log$ scale). 
		%Note that four points have deviations of the order of $10^{-3}$ and are not visible on the figure.
		}
		\label{fig:Analyze_V_points}
	\end{figure}

	Upon completion, the loss function converges around $0.002$ \gl{$10^{-3}$?}. The distribution of the values of the gradient at each of the data points is plotted in fig.~\ref{fig:log_grad_V}. More than $93\%$ of the points converged to values of $\vert\nabla V\vert$ lower than $10^{-4}$, and more than $99\%$ of them exhibited gradient norms smaller than $10^{-3}$, yielding a satisfying sampling of the flat directions. As can be observed in fig.~\ref{fig:log_Vp4}, all data points converged to values close to $V(0)=-4$. Specifically, nearly all points attained $V(0)$ within an absolute error of at least $10^{-4}$, with the exception of four points whose deviations were of the order of $10^{-3}$ (not visible on the graph). As the origin $\vec{X}=0$ of the parameter space corresponds to the three-dimensional truncation of the round ${\rm AdS}_{3}\times S^{3}$ solution of half-maximal supergravity in six dimensions with $V(0)=-4$, this confirms that the data points lie within flat directions of the potential.

@@

	As a first visualisation, we present a tomography of the data in fig.~\ref{triangular_plot_1_2_4_8_10}. This figure shows all possible 2d projections of the data, along with the 1d histograms of each coordinate after gradient descent. Note that all directions appear to be well populated. There also seem to be non-trivial correlations in the data, see for example the $x_1/x_2$, $x_4/\tilde{x}_8$ or $x_4/x_{10}$ graphs. They could be genuine correlations, or may result from larger basins of attraction. Thanks to the subsequent analysis, we find that the structures in the $x_1/x_2$ plot are artefacts of the gradient descent, whereas the ones seen in the $x_4/\tilde{x}_8$ or $x_4/x_{10}$ plots reflect genuine features of the manifold.
